---
title: "KELPER"
subtitle: "Kelp demography data"
author: "Tim Szewczyk"
output:
  html_document:
    theme: spacelab
    df_print: paged
    anchor_sections: TRUE
    toc: yes
    toc_depth: 2
    toc_float: true
  pdf_document:
    toc: yes
  html_notebook:
    theme: spacelab
    toc: yes
editor_options: 
  chunk_output_type: console
---

# Overview

This file explores demographic data, focussing on digitized plots from Kain on *Laminaria hyperborea*. 


```{r setup, echo=F, include=F}
# set directory for knitr as main project directory
knitr::opts_knit$set(root.dir=rprojroot::find_rstudio_root_file())
```

```{r workspace, echo=F, include=F, message=F, warning=F}
library(tidyverse); library(brms); theme_set(theme_bw())
source("code/002_fnDemography.R")

digitizedMetadata <- read_csv("..\\data\\digitized\\metadata.csv", show_col_types=F)

fig.f <- unique(str_remove(dir("..\\data\\digitized", "_fig.*csv"), "[a-z]?.csv"))

digitized <- fig.f %>% 
  map(~dir("..\\data\\digitized", paste0(.x, "[a-z]?.csv")) %>%
        map_dfr(~read_csv(paste0("..\\data\\digitized\\", .x), show_col_types=F) %>%
                  mutate(reference=str_remove(.x, ".csv"))) %>%
        left_join(., digitizedMetadata, by="reference")) %>%
  setNames(fig.f)

fig_var <- imap_dfr(digitized, 
                    ~tibble(V1=names(.x)[1], V2=names(.x)[2], reference=.y)) %>%
  arrange(V1, V2) %>%
  mutate(vars=paste0(V1, "_", V2))

unique_comparisons <- unique(fig_var$vars)
varCombined.ls <- vector("list", length(unique_comparisons)) %>% 
  setNames(unique_comparisons)
for(i in seq_along(varCombined.ls)) {
  refs <- fig_var$reference[fig_var$vars==unique_comparisons[i]]
  varCombined.ls[[i]] <- do.call('rbind', digitized[refs])
}
```



# Growth





# Allometry
Frond weight generally increases linearly with stipe weight on a log-log scale, though plants seem to be frondier in new growth after a clearing (Kain 1963, 1976a). 
```{r wtStipe_wtFrond_clearing, echo=F}
varCombined.ls$weightStipe_weightFrond %>%
  ggplot(aes(log(weightStipe), log(weightFrond), colour=habitat=="clearing")) +
  geom_point() + stat_smooth(method="lm") + 
  scale_colour_brewer(type="qual", palette=3)
```

The effect of depth is less clear when aggregating across studies, though within Kain 1963, shallower kelp is frondier (blue vs. yellow).
```{r wtStipe_wtFrond_clearing}
varCombined.ls$weightStipe_weightFrond %>% 
  ggplot(aes(log(weightStipe), log(weightFrond), 
             colour=depth, group=depth, shape=habitat=="clearing")) +
  geom_point() + stat_smooth(method="lm") + 
  scale_shape_manual("Clearing", values=c(16, 1)) +
  scale_colour_viridis_c(na.value="black")
```




# Density





# Survival




