---
title: "KELPER"
subtitle: "Kelp demography data"
author: "Tim Szewczyk"
output:
  html_document:
    theme: spacelab
    df_print: paged
    anchor_sections: TRUE
    toc: yes
    toc_depth: 2
    toc_float: true
  pdf_document:
    toc: yes
  html_notebook:
    theme: spacelab
    toc: yes
editor_options: 
  chunk_output_type: console
---

# Overview

This file explores demographic data, focussing on digitized plots from Kain on *Laminaria hyperborea*. 

So far, the data we have allows us to:

- Convert between stipe length (growth, stage assignment), frond weight (harvestable biomass), and FAI (light-based density effects), with the conversion to FAI contingent on depth   
- Estimate maximum growth rates of frond length, frond weight, and FAI by date
- Estimate maximum densities based on FAI, contingent on depth
- Estimate maximum stipe length, contingent on depth


We are still **missing**:

- Survival by size
- Effect of density on growth rate
- Effect of density on survival rate
- Effects of environmental covariates
  - SST (growth rate, survival rate, max length, allometry)
  - chl a (growth rate, survival rate, max density, allometry)
  - wave fetch (growth rate, survival rate, max length, allometry)


```{r setup, echo=F, include=F}
# set directory for knitr as main project directory
knitr::opts_knit$set(root.dir=rprojroot::find_rstudio_root_file())
```

```{r workspace, echo=F, include=F, message=F, warning=F}
library(tidyverse); library(sf); library(brms); theme_set(theme_bw())
source("code/000_fnMisc.R"); source("code/002_fnDemography.R")

data.dir <- "..\\data\\digitized\\"
data.ls <- compileDatasets(data.dir)

site.i <- read_csv(paste0(data.dir, "sitesDigitized.csv"), col_select=1:3, show_col_types=F)
site.sf <- site.i %>% filter(!is.na(lon) & !is.na(lat)) %>%
  st_as_sf(coords=c("lon", "lat")) %>% st_set_crs(4326)
coast <- read_sf("../../0_gis/dataGeog/ne_10m_land.shp") %>%
  add_row(read_sf("../../0_gis/dataGeog/ne_10m_minor_islands.shp")) %>%
  st_crop(st_buffer(site.sf, 5e4))
```



```{r maps, echo=F, warning=F, message=F}
ggplot(site.sf) + geom_sf(data=coast) + geom_sf(colour="red")
  
```







# Allometry

We are interested in three main morphological metrics: stipe length (stage assignment, tracking growth between timesteps), frond weight (harvestable biomass), and frond area index (FAI: frond area / substrate area). 

The conversion between stipe length and frond weight depends on light availability, with somewhat more allocation to fronds with abundant light, and somewhat leggier plants when light is scarce. This is inferred from Kain (1963, 1976a) based on plants in recovering clearings compared to plants in more natural patches.

```{r wtStipe_wtFrond_clearing, echo=F}
data.ls$weightStipe_weightFrond %>%
  ggplot(aes(log(weightStipe), log(weightFrond), colour=habitat)) +
  geom_point() + stat_smooth(method="lm") + 
  scale_colour_brewer(type="qual", palette=3)
```

The effect of depth is less clear. Within habitat types, shallower kelp is frondier (blue vs. yellow), consistent with the effect of light availability. However, the shallowest site is intermediate, perhaps reflecting the effect of the forest habitat compared to the patch limit. The effect of clearing remains.

```{r wtStipe_wtFrond_habitatDepth, echo=F}
data.ls$weightStipe_weightFrond %>% 
  ggplot(aes(log(weightStipe), log(weightFrond), 
             colour=-depth, group=depth, shape=habitat, linetype=habitat)) +
  geom_point() + stat_smooth(method="lm") + 
  scale_shape_manual(values=c(16, 2, 1)) +
  scale_colour_viridis_c(na.value="black")
```

One possibility would be to use a weighted average (i.e., calculate regression coefficients using habitat=="clearing" and habitat!="clearing"), then use a weighted average based on the harvested area.

```{r lm_lenStipe_wtStipe_wtFrond}
lenStipe_wtStipe <- brm(log(weightStipe) ~ log(lengthStipe),
                        data=data.ls$lengthStipe_weightStipe)
wtStipe_wtFrond.clearing <- data.ls$weightStipe_weightFrond %>%
  filter(habitat=="clearing") %>%
  brm(log(weightFrond) ~ log(weightStipe), data=.)
wtStipe_wtFrond.nonclearing <- data.ls$weightStipe_weightFrond %>%
  filter(habitat!="clearing") %>%
  brm(log(weightFrond) ~ log(weightStipe), data=.)

plot(NA, NA, xlim=c(log(0.06), log(475)), ylim=c(log(0.4), log(1800)), 
     xlab="Stipe weight", ylab="Frond weight", main="Red = more clearing")
clear.prop <- seq(0, 1, by=0.1)
coef.clear <- summary(wtStipe_wtFrond.clearing)$fixed[,1]
coef.nonclear <- summary(wtStipe_wtFrond.nonclearing)$fixed[,1]
for(i in seq_along(clear.prop)) { 
  abline(coef.nonclear[1]*(1-clear.prop[i])+coef.clear[1]*clear.prop[i],
         coef.nonclear[2]*(1-clear.prop[i])+coef.clear[2]*clear.prop[i], 
         col=rgb(clear.prop[i],0,0))
}
```

Then, the conversion from frond weight to frond area explicitly accounts for depth, as the relationship is clearly depth dependent. As it appears to be log-linear, we can estimate the change in slope across depths. For a given depth, we can then calculate the frond area based on the frond weight, accounting for uncertainty in the relationship.
```{r wtFrond_areaFrond}
data.ls$depth_beta_mn %>%
  ggplot(aes(depth, y=beta_mn, ymin=beta_mn-beta_sd, ymax=beta_mn+beta_sd)) +
  geom_pointrange() + 
  labs(x="Depth (m)", y="Slope of frond log area ~ log weight")

beta.sim <- data.ls$depth_beta_mn %>%
  select(depth, beta_mn, beta_sd) %>%
  expand_grid(sim=1:100, .) %>%
  mutate(value=rnorm(n(), beta_mn, beta_sd))
depth_wtareaFrondBeta <- brm(value ~ log(depth), data=beta.sim)
```

Then, for a given depth and level of clearing, we can convert from stipe length to frond area using the appropriate regression coefficients and incorporating uncertainty by calculating across each sample from the posterior. If we use only two or three size classes, this will be a big simplification, but maybe that's ok. I think it's necessary.




# Growth

Younger plants show faster stipe growth, but older plants show faster frond growth.

Stipe growth is seasonal, but less clearly than frond growth. Most growth occurs between February and July, but with a bit more in December and January, and less clear coordination among age groups. Younger (presumably shorter) plants show greater growth. 
```{r growRateFrondLength, echo=F}
data.ls$day_growthStipe %>%
  mutate(dateMeasured=lubridate::as_date(round(day))) %>%
  ggplot(aes(dateMeasured, growthStipe/7, colour=factor(ageGroup))) + 
  geom_line(size=1) + scale_colour_viridis_d("Age") +
  labs(x="Date", y="Stipe growth (mm/day)") +
  scale_x_date(date_breaks="1 month", date_labels="%b", 
               limits=lubridate::dmy(c("01-01-1970", "01-01-1971"))) +
  theme(panel.grid.minor.x=element_blank())

data.ls$day_growthStipe %>% select(-location, -depth, -age, -date, -lat, -lon) %>%
  arrange(ageGroup, day) %>% 
  mutate(dateMeasured=lubridate::as_date(round(day))) %>%
  group_by(ageGroup) %>%
  mutate(datePrevious=lag(dateMeasured),
         nDays=dateMeasured-datePrevious,
         month=lubridate::month(dateMeasured)) %>%
  ggplot(aes(factor(month), growthStipe/7, colour=ageGroup<4)) + geom_point()

library(geoTS)
recruits <- data.ls$day_growthStipe %>% filter(ageGroup < 2) %>%
  mutate(dailyGrowth=growthStipe/7)
subcanopy <- data.ls$day_growthStipe %>% filter(ageGroup > 0 & ageGroup < 3) %>%
  mutate(dailyGrowth=growthStipe/7)
canopy <- data.ls$day_growthStipe %>% filter(ageGroup > 2) %>%
  mutate(dailyGrowth=growthStipe/7)
recruits.hm <- haRmonics(y=recruits$dailyGrowth,
                         ts=recruits$day,
                         method="hants",
                          low=-20, high=20, fitErrorTol=0.5, degreeOverDeter=1,
                         lenBasePeriod=365, numFreq=1, delta=0.1)
subcanopy.hm <- haRmonics(y=subcanopy$dailyGrowth,
                          ts=subcanopy$day,
                          method="hants",
                          low=-20, high=20, fitErrorTol=0.5, degreeOverDeter=10,
                          lenBasePeriod=365, numFreq=1, delta=0.1)
canopy.hm <- haRmonics(y=canopy$dailyGrowth,
                       ts=canopy$day,
                       method="hants",
                          low=-20, high=20, fitErrorTol=0.5, degreeOverDeter=10,
                       lenBasePeriod=365, numFreq=1, delta=0.1)


hm.1 <- list(a=recruits.hm$a.coef, b=recruits.hm$b.coef, 
             fit=recruits.hm$fitted, 
             phase=recruits.hm$phase, amp=recruits.hm$amplitude)
hm.2 <- list(a=subcanopy.hm$a.coef, b=subcanopy.hm$b.coef, 
             fit=subcanopy.hm$fitted, 
             phase=subcanopy.hm$phase, amp=subcanopy.hm$amplitude)
hm.3 <- list(a=canopy.hm$a.coef, b=canopy.hm$b.coef, 
             fit=canopy.hm$fitted, 
             phase=canopy.hm$phase, amp=canopy.hm$amplitude)

library(lubridate)

start <- ymd("1970-01-01")
end <- ymd("1970-12-31")

day_seq <- seq(start, end, by="day")
y <- (yday(day_seq)-1)/c(364,365)[leap_year(day_seq)+1]
pred.1 <- mean(hm.1$fit) + hm.1$a * cos(2*pi*y) + hm.1$b * sin(2*pi*y)
pred.2 <- mean(hm.2$fit) + hm.2$a * cos(2*pi*y) + hm.2$b * sin(2*pi*y)
pred.3 <- mean(hm.3$fit) + hm.3$a * cos(2*pi*y) + hm.3$b * sin(2*pi*y)

plot(NA, NA, xlim=c(1,length(y)), ylim=c(0, 1.5), xlab="Day", ylab="Daily growth")
lines(1:length(y), pred.1, col="blue")
points(recruits$day, recruits$dailyGrowth, col="blue")
lines(1:length(y), pred.2, col="red")
points(subcanopy$day, subcanopy$dailyGrowth, col="red")
lines(1:length(y), pred.3, col="gray")
points(canopy$day, canopy$dailyGrowth, col="gray")

tibble(day=day_seq, 
       month=month(day_seq),
       season=case_when(month %in% 2:6 ~ "spring",
                        month %in% 7:9 ~ "fall",
                        month %in% c(10:12,1) ~ "winter"),
       growthRecuits=pred.1,
       growthSubcanopy=pred.2,
       growthCanopy=pred.3) %>% 
  group_by(season) %>%
  summarise(across(contains("growth"), sum))
```



Frond growth is highly seasonal, with essentially all growth occurring between February and July, and a peak in April. For the length of the central frond finger, growth rates are comparable for 1-3 year olds, then somewhat lower for 4 year olds.
```{r growRateFrondLengthPlot, echo=F}
data.ls$day_growthFrond %>%
  mutate(dateMeasured=lubridate::as_date(round(day))) %>%
  ggplot(aes(dateMeasured, growthFrond/7, colour=factor(ageGroup))) + 
  geom_line(size=1) + scale_colour_viridis_d("Age") +
  labs(x="Date", y="Frond growth (mm/day)") +
  scale_x_date(date_breaks="1 month", date_labels="%b", 
               limits=lubridate::dmy(c("01-01-1970", "01-01-1971"))) +
  theme(panel.grid.minor.x=element_blank())
```


For frond area, however, growth rate increases from ages 1-3, with similar seasonality across age groups. The peak occurs in May.
```{r growRateFrondArea, echo=F}
data.ls$day_areaFrond %>%
  group_by(age, location) %>%
  mutate(deltaArea=areaFrond-lag(areaFrond),
         deltaDay=day-lag(day),
         rate=deltaArea/deltaDay,
         midptDay=day+deltaDay/2) %>%
  mutate(dateMeasured=lubridate::as_date(round(midptDay))) %>%
  ggplot(aes(dateMeasured, rate, colour=factor(age), linetype=location)) + 
  geom_line(size=1) + scale_colour_viridis_d("Age") +
  labs(x="Date", y=expression('Growth rate in frond area (cm'^2~'/day)')) +
  theme(panel.grid.minor.x=element_blank())
```


```{r growRateFAI, echo=F}
data.ls$day_FAI %>%
  mutate(habitat=str_remove(habitat, "clearing - "),
         clearanceDate=lubridate::ymd(paste0("1970-", habitat, "-01")),
         dateMeasured=clearanceDate+day) %>%
  ggplot(aes(dateMeasured, FAI, colour=habitat)) + 
  geom_point() + 
  scale_colour_viridis_d("Month of\nclearance") +
  labs(x="Days post-clearance", y="Frond Area Index") +
  theme(panel.grid.minor.x=element_blank())
```






# Survival

```{r}
data.ls$stageFrom_stageTo %>%
  mutate(type=case_when(stageTo=="dead" ~ "mortality",
                        stageFrom==stageTo ~ "persist",
                        stageFrom=="canopy" & stageTo=="recruits" ~ "fecundity",
                        stageFrom=="subcanopy" & stageTo=="canopy" ~ "growth",
                        stageFrom=="recruits" & stageTo=="subcanopy" ~ "growth"),
         exposure=factor(exposure, levels=c("low", "medium", "high")),
         stageFrom=factor(stageFrom, levels=c("canopy", "subcanopy", "recruits"))) %>%
  ggplot(aes(as.numeric(exposure), rate, colour=stageFrom)) + 
  geom_line(size=1) + facet_wrap(~type, scales="free_y") +
  scale_colour_brewer(type="qual", palette=2)
```













# Density

**These numbers are huge and lead to absurd FAI. Use Kain 1977 Fig 12-13 instead**
**I changed the code, but not the text**
Kain (1976a) reports kelp densities by stipe length in virgin forests and following recovery from a clearing event, with the latter divided by age. Data are from the Isle of Man as well as a previous publication from Norway. 

Density in the clearings is much higher, with a distribution skewed toward shorter, presumably younger, plants. As kelp is expected to self-thin, this is not surprising. The Norwegian virgin forest appears to have no small plants, while the Scottish forest is vaguely similar in shape, but muted in amplitude.  
```{r density_lenStipe, echo=F, message=F}
data.ls$lengthStipe_NperSqM <- data.ls$lengthStipe_NperSqM %>%
  filter(!grepl("fig17", reference))
data.ls$lengthStipe_NperSqM %>%
  ggplot(aes(lengthStipe, ymax=NperSqM, fill=depth, group=reference)) + 
  geom_ribbon(ymin=0, alpha=0.5, colour="grey30") + 
  # facet_grid(location~.) +
  scale_fill_viridis_c() 

data.ls$lengthStipe_NperSqM %>% 
  group_by(location, depth) %>%
  summarise(NperSqM=sum(NperSqM)) %>%
  ggplot(aes(depth, NperSqM)) + 
  geom_point() + ylim(0, NA) +
  scale_fill_viridis_c() 

data.ls$lengthStipe_NperSqM %>% 
  mutate(lengthBin=lengthStipe %/% 100 * 100 + 50) %>%
  group_by(location, depth, lengthBin) %>%
  summarise(NperSqM=sum(NperSqM)) %>%
  ggplot(aes(lengthBin, NperSqM, colour=depth, group=paste(location,depth))) + 
  geom_point() + geom_line(size=1) +
  scale_colour_viridis_c() 
```

As there seems to be a natural break in the datasets around 600mm. Using this as a threshold for *canopy* vs. *subcanopy* plants (the hoped-for stages), we can estimate the density of individuals in each category.
```{r density_depth, echo=F, message=F}
data.ls$lengthStipe_NperSqM %>% 
  mutate(stage=c("canopy", "subcanopy")[1 + (lengthStipe < 600)]) %>%
  group_by(location, depth, stage) %>%
  summarise(NperSqM=sum(NperSqM)) %>%
  ggplot(aes(depth, NperSqM, colour=stage, group=stage)) + 
  geom_point() + stat_smooth(method="lm") +
  scale_y_log10() +
  scale_colour_brewer(type="qual", palette=2)
```

Or, following Kain 1977 (fig. 14, LAI vs. stipe length distribution), we can use the upper 1/3 of the length distribution as a threshold for *canopy* vs. *subcanopy* plants (the hoped-for stages).
```{r density_canopy, echo=F, message=F}
stipe.rng <- data.ls$lengthStipe_NperSqM %>% 
  group_by(depth) %>% 
  summarise(stipeMin=min(lengthStipe), stipeMax=max(lengthStipe), 
            top90=(stipeMax-stipeMin)*.1 + stipeMin,
            top75=(stipeMax-stipeMin)*.25 + stipeMin,
            top66=(stipeMax-stipeMin)*1/3 + stipeMin,
            top50=(stipeMax-stipeMin)*.5 + stipeMin,
            top33=(stipeMax-stipeMin)*2/3 + stipeMin,
            top25=(stipeMax-stipeMin)*.75 + stipeMin,
            top10=(stipeMax-stipeMin)*.9 + stipeMin)
data.ls$lengthStipe_NperSqM %>% 
  left_join(., stipe.rng, by="depth") %>%
  mutate(stage=c("canopy", "subcanopy")[1 + (lengthStipe < top33)]) %>%
  group_by(depth, stage) %>%
  summarise(NperSqM=sum(NperSqM)) %>%
  ggplot(aes(stage, NperSqM, fill=depth, group=depth)) + 
  geom_bar(stat="identity", position="dodge", colour="grey30") +
  scale_fill_viridis_c() 
data.ls$lengthStipe_NperSqM %>% 
  left_join(., stipe.rng, by="depth") %>%
  mutate(stage=c("canopy", "subcanopy")[1 + (lengthStipe < top33)]) %>%
  group_by(depth, stage) %>%
  summarise(NperSqM=sum(NperSqM)) %>%
  ggplot(aes(stage, NperSqM)) + 
  geom_boxplot() 
data.ls$lengthStipe_NperSqM %>% 
  left_join(., stipe.rng, by="depth") %>%
  mutate(stage=c("canopy", "subcanopy")[1 + (lengthStipe < top33)]) %>%
  group_by(depth, stage) %>%
  summarise(NperSqM=sum(NperSqM)) %>%
  ggplot(aes(depth, NperSqM, colour=stage)) + 
  geom_point() + stat_smooth(method="lm") +
  scale_y_log10() + scale_x_log10()
```


Alternatively, we can use allometric relationships above to convert from stipe length to frond weight. This is useful as harvesting biomass, but we really need to get from frond weight to frond area index.
```{r densitylenStipe_wtFrond, eval=FALSE, include=FALSE, results=F}
lenStipe_wtStipe <- brm(log(weightStipe) ~ log(lengthStipe),
                        data=data.ls$lengthStipe_weightStipe)
wtStipe_wtFrond <- brm(log(weightFrond) ~ log(weightStipe), 
                       data=data.ls$weightStipe_weightFrond)
lenStipe_wtStipe.mx <- as_draws_matrix(lenStipe_wtStipe)
wtStipe_wtFrond.mx <- as_draws_matrix(wtStipe_wtFrond)


simStipeLen <- data.ls$lengthStipe_NperSqM$lengthStipe %/% 100 * 100 + 50
estFrondWt <- estStipeWt <- matrix(0, nrow(wtStipe_wtFrond.mx), length(simStipeLen))
for(i in 1:nrow(wtStipe_wtFrond.mx)) {
  estStipeWt[i,] <- convertAllometry(log(simStipeLen), lenStipe_wtStipe.mx[i,1:2],
                                   shape="linear", 
                                   resid_err=lenStipe_wtStipe.mx[i,3])
  estFrondWt[i,] <- convertAllometry(estStipeWt[i,], wtStipe_wtFrond.mx[i,1:2],
                                       shape="linear", 
                                       resid_err=wtStipe_wtFrond.mx[i,3])
}
data.ls$lengthStipe_NperSqM <- data.ls$lengthStipe_NperSqM %>%
  mutate(wtStipe_mn=apply(exp(estStipeWt), 2, mean),
         wtStipe_q10=apply(exp(estStipeWt), 2, quantile, probs=0.1),
         wtStipe_q90=apply(exp(estStipeWt), 2, quantile, probs=0.9),
         wtFrond_mn=apply(exp(estFrondWt), 2, mean),
         wtFrond_q10=apply(exp(estFrondWt), 2, quantile, probs=0.1),
         wtFrond_q90=apply(exp(estFrondWt), 2, quantile, probs=0.9))

```

```{r density_wtFrond, message=F, eval=F}
data.ls$lengthStipe_NperSqM %>% 
  mutate(lengthBin=lengthStipe %/% 100 * 100 + 50) %>%
  group_by(location, depth, lengthBin) %>%
  summarise(wtStipePerSqM_mn=sum(wtStipe_mn*NperSqM)/1000,
            wtStipePerSqM_q10=sum(wtStipe_q10*NperSqM)/1000,
            wtStipePerSqM_q90=sum(wtStipe_q90*NperSqM)/1000) %>%
  ggplot(aes(lengthBin, wtStipePerSqM_mn, colour=depth, fill=depth, group=depth)) + 
  geom_ribbon(aes(ymin=wtStipePerSqM_q10, ymax=wtStipePerSqM_q90), alpha=0.5) +
  geom_point() + 
  scale_colour_viridis_c() + scale_fill_viridis_c() +
  facet_grid(location~.) +
  labs(x="Stipe length (mm)", y=expression('Stipe weight (kg/m'^2~')'))

data.ls$lengthStipe_NperSqM %>% 
  left_join(., stipe.rng, by="depth") %>%
  mutate(stage=c("canopy", "subcanopy")[1 + (lengthStipe < top33)]) %>%
  filter(stage=="canopy") %>%
  group_by(location, depth, stage) %>%
  summarise(wtStipePerSqM_mn=sum(wtStipe_mn*NperSqM)/1e3,
            wtStipePerSqM_q10=sum(wtStipe_q10*NperSqM)/1e3,
            wtStipePerSqM_q90=sum(wtStipe_q90*NperSqM)/1e3) %>%
  ggplot(aes(depth, wtStipePerSqM_mn, colour=depth, group=depth)) + 
  geom_pointrange(aes(ymin=wtStipePerSqM_q10, ymax=wtStipePerSqM_q90), 
                  position=position_dodge(width=0.25)) +
  scale_colour_viridis_c() + scale_y_log10() +
  labs(x="", y=expression('Stipe weight (kg/m'^2~')'))


data.ls$lengthStipe_NperSqM %>% 
  mutate(lengthBin=lengthStipe %/% 100 * 100 + 50) %>%
  group_by(location, depth, lengthBin) %>%
  summarise(wtFrondPerSqM_mn=sum(wtFrond_mn*NperSqM)/1000,
            wtFrondPerSqM_q10=sum(wtFrond_q10*NperSqM)/1000,
            wtFrondPerSqM_q90=sum(wtFrond_q90*NperSqM)/1000) %>%
  ggplot(aes(lengthBin, wtFrondPerSqM_mn, colour=depth, fill=depth, group=depth)) + 
  geom_ribbon(aes(ymin=wtFrondPerSqM_q10, ymax=wtFrondPerSqM_q90), alpha=0.5) +
  geom_point() + 
  scale_colour_viridis_c() + scale_fill_viridis_c() +
  facet_grid(location~.) +
  labs(x="Stipe length (mm)", y=expression('Frond weight (kg/m'^2~')'))

data.ls$lengthStipe_NperSqM %>% 
  group_by(location, depth) %>%
  summarise(wtFrondPerSqM_mn=sum(wtFrond_mn*NperSqM)/1e3,
            wtFrondPerSqM_q10=sum(wtFrond_q10*NperSqM)/1e3,
            wtFrondPerSqM_q90=sum(wtFrond_q90*NperSqM)/1e3) %>%
  ggplot(aes(depth, wtFrondPerSqM_mn, colour=depth, shape=location)) + 
  geom_pointrange(aes(ymin=wtFrondPerSqM_q10, ymax=wtFrondPerSqM_q90), 
                  position=position_dodge(width=0.25)) +
  scale_colour_viridis_c() +
  labs(x="Depth (m)", y=expression('Frond weight (kg/m'^2~')'))

data.ls$lengthStipe_NperSqM %>% 
  mutate(stage=c("canopy", "subcanopy")[1 + (lengthStipe < 600)]) %>%
  group_by(location, depth, stage) %>%
  summarise(wtFrondPerSqM_mn=sum(wtFrond_mn*NperSqM)/1e3,
            wtFrondPerSqM_q10=sum(wtFrond_q10*NperSqM)/1e3,
            wtFrondPerSqM_q90=sum(wtFrond_q90*NperSqM)/1e3) %>%
  ggplot(aes(stage, wtFrondPerSqM_mn, colour=depth, group=depth, shape=location)) + 
  geom_pointrange(aes(ymin=wtFrondPerSqM_q10, ymax=wtFrondPerSqM_q90), 
                  position=position_dodge(width=0.25)) +
  scale_colour_viridis_c() +
  labs(x="", y=expression('Frond weight (kg/m'^2~')'))

data.ls$lengthStipe_NperSqM %>% 
  left_join(., stipe.rng, by="depth") %>%
  mutate(stage=c("canopy", "subcanopy")[1 + (lengthStipe < top33)]) %>%
  filter(stage=="canopy") %>%
  group_by(location, depth, stage) %>%
  summarise(wtFrondPerSqM_mn=sum(wtFrond_mn*NperSqM)/1e3,
            wtFrondPerSqM_q10=sum(wtFrond_q10*NperSqM)/1e3,
            wtFrondPerSqM_q90=sum(wtFrond_q90*NperSqM)/1e3) %>%
  ggplot(aes(depth, wtFrondPerSqM_mn, colour=depth, group=depth)) + 
  geom_pointrange(aes(ymin=wtFrondPerSqM_q10, ymax=wtFrondPerSqM_q90), 
                  position=position_dodge(width=0.25)) +
  scale_colour_viridis_c() +
  labs(x="", y=expression('Frond weight (kg/m'^2~')'))
```

And further, we can estimate frond area from the frond weight. **Depth should be added here.**
```{r densitylenStipe_areaFrond, results=F, eval=F}
lenStipe_wtStipe <- brm(log(weightStipe) ~ log(lengthStipe),
                        data=data.ls$lengthStipe_weightStipe)
wtStipe_wtFrond <- brm(log(weightFrond) ~ log(weightStipe), 
                       data=data.ls$weightStipe_weightFrond)
wtFrond_areaFrond <- brm(log(areaFrond/1e4) ~ log(weightFrond), 
                       data=data.ls$weightFrond_areaFrond)
lenStipe_wtStipe.mx <- as_draws_matrix(lenStipe_wtStipe)
wtStipe_wtFrond.mx <- as_draws_matrix(wtStipe_wtFrond)
wtFrond_areaFrond.mx <- as_draws_matrix(wtFrond_areaFrond)


simStipeLen <- data.ls$lengthStipe_NperSqM$lengthStipe %/% 100 * 100
estFrondArea <- matrix(0, nrow(wtFrond_areaFrond.mx), length(simStipeLen))
for(i in 1:nrow(wtFrond_areaFrond.mx)) {
  estStipeWt <- convertAllometry(log(simStipeLen), lenStipe_wtStipe.mx[i,1:2],
                                   shape="linear")#, 
                                   # resid_err=lenStipe_wtStipe.mx[i,3])
  estFrondWt <- convertAllometry(estStipeWt, wtStipe_wtFrond.mx[i,1:2],
                                   shape="linear")#, 
                                   # resid_err=wtStipe_wtFrond.mx[i,3])
  estFrondArea[i,] <- convertAllometry(estFrondWt, wtFrond_areaFrond.mx[i,1:2],
                                         shape="linear")#, 
                                         # resid_err=wtFrond_areaFrond.mx[i,3])
}
data.ls$lengthStipe_NperSqM <- data.ls$lengthStipe_NperSqM %>%
  mutate(areaFrond_mn=apply(exp(estFrondArea), 2, mean),
         areaFrond_q10=apply(exp(estFrondArea), 2, quantile, probs=0.1),
         areaFrond_q90=apply(exp(estFrondArea), 2, quantile, probs=0.9))
```

```{r density_areaFrond, message=F, eval=F}
data.ls$lengthStipe_NperSqM %>% 
  mutate(lengthBin=lengthStipe %/% 100 * 100 + 50) %>%
  group_by(location, depth, lengthBin) %>%
  summarise(FAI_mn=sum(areaFrond_mn*NperSqM),
            FAI_q10=sum(areaFrond_q10*NperSqM),
            FAI_q90=sum(areaFrond_q90*NperSqM)) %>%
  ggplot(aes(lengthBin, FAI_mn, colour=depth, fill=depth, group=depth)) + 
  geom_ribbon(aes(ymin=FAI_q10, ymax=FAI_q90), alpha=0.5) +
  geom_point() + 
  scale_colour_viridis_c() + scale_fill_viridis_c() +
  facet_grid(location~.) +
  labs(x="Stipe length (mm)", y="Frond Area Index")

data.ls$lengthStipe_NperSqM %>% 
  group_by(location, depth) %>%
  summarise(FAI_mn=sum(areaFrond_mn*NperSqM),
            FAI_q10=sum(areaFrond_q10*NperSqM),
            FAI_q90=sum(areaFrond_q90*NperSqM)) %>%
  ungroup %>% mutate(source="modelled") %>%
  add_row(data.ls$depth_FAI %>% select(location, depth, FAI) %>%
            rename(FAI_mn=FAI) %>% mutate(source="data")) %>%
  filter(FAI_mn < 6) %>%
  ggplot(aes(depth, FAI_mn, colour=depth, shape=source, linetype=source)) + 
  geom_pointrange(aes(ymin=FAI_q10, ymax=FAI_q90)) +
  stat_smooth(method="lm", colour="grey30", se=F, size=0.6) +
  scale_colour_viridis_c() + scale_shape_manual(values=c(16, 1)) +
  labs(x="Depth (m)", y="Frond Area Index")

data.ls$lengthStipe_NperSqM %>% 
  left_join(., stipe.rng, by="depth") %>%
  mutate(stage=case_when(lengthStipe<top66 ~ "sub-subcanopy",
                         lengthStipe>=top66 & lengthStipe<top33 ~ "subcanopy",
                         lengthStipe>=top33 ~ "canopy")) %>%
  mutate(stage=factor(stage, levels=c("sub-subcanopy", "subcanopy", "canopy"))) %>%
  group_by(location, depth, stage) %>%
  summarise(FAI_mn=sum(areaFrond_mn*NperSqM),
            FAI_q10=sum(areaFrond_q10*NperSqM),
            FAI_q90=sum(areaFrond_q90*NperSqM)) %>%
  ggplot(aes(depth, FAI_mn, colour=stage)) + 
  geom_pointrange(aes(ymin=FAI_q10, ymax=FAI_q90), 
                  position=position_dodge(width=0.25)) +
  stat_smooth(method="lm", se=F, size=0.6) +
  scale_colour_brewer(type="qual", palette=2) +
  labs(x="Depth (m)", y="Frond Area Index")
```








# Size distribution

The size distribution, measured by stipe length, is qualitatively less bimodal in sparsely populated patches compared to within forests. 
```{r lenStipe_distribution, echo=F}
data.ls$lengthStipe_proportion %>% filter(!grepl("fig14a", reference)) %>%
  ggplot(aes(lengthStipe, ymax=proportion, fill=location, group=reference)) + 
  geom_ribbon(ymin=0, alpha=0.5) + facet_grid(depth~.)
```




# Depth
Stipe weight levels off after age 5-6, and individuals tend to have more massive stipes at shallower depths, in forests, or both.

```{r wtStipe_age_depth, echo=F}
data.ls$ageMin_weightStipe %>%
  ggplot(aes(factor(ageMin), log(weightStipe), fill=factor(depth))) + 
  geom_boxplot() + 
  labs(x="Age", y="ln stipe weight (g)") +
  scale_fill_brewer("Location")

data.ls$ageMin_weightStipe %>%
  ggplot(aes(ageMin, log(weightStipe), colour=factor(depth))) + 
  geom_point(alpha=0.5) + stat_smooth(method="lm", formula=y~log(x)) +
  labs(x="Age", y="ln stipe weight (g)") +
  scale_colour_brewer("Depth", type="qual", palette=2)
```

